Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections

### 2022-05-27 manuscript.tex
deploy movies [ci skip]
This build is based on
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/26bf17f811e80363e8f38656b89bbae2c14147f8.

This commit was created by the following CI build and job:
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/commit/26bf17f811e80363e8f38656b89bbae2c14147f8/checks
https://github.com/jojoelfe/deco_lace_template_matching_manuscript/actions/runs/2398856270

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections},
  pdfauthor={Johannes ElferichNikolaus Grigorieff},
  pdflang={en-US},
  pdfkeywords={cryo-EM, visual protemics, ribosome},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\usepackage[margin=2cm]{geometry}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage{cleveref}

@@

This manuscript
(\href{https://jojoelfe.github.io/deco_lace_template_matching_manuscript/v/e94f17b00fe53a7b703becda30c7a1b1503e6e7e/}{permalink})
was automatically generated
from \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/tree/e94f17b00fe53a7b703becda30c7a1b1503e6e7e}{jojoelfe/deco\_lace\_template\_matching\_manuscript@e94f17b}
on May 20, 2022.

@@

\begin{itemize}
\item
  \textbf{Johannes Elferich}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9911-706X}{0000-0002-9911-706X}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.pdf}
  \href{https://github.com/jojoelfe}{jojoelfe}
  RNA Therapeutic Institute, UMass Chan Medical School; HHMI
\item
  \textbf{Nikolaus Grigorieff}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-1506-909X}{0000-0002-1506-909X}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.pdf}
  \href{https://github.com/nikogrigorieff}{nikogrigorieff}
  RNA Therapeutic Institute, UMass Chan Medical School; HHMI
\end{itemize}

@@

@@

Grids were loaded into a Titan Krios TEM (Thermo Fisher) operated at 300 keV and
equipped with a BioQuantum energy filter (Gatan) and K3 camera (Gatan). The
microscope was aligned using a cross-grating grid on the stage. Prior to each
session, we carefully performed the ``Image/Beam'' calibration in nanoprobe. We
set the magnification to a pixel size of 1.76 Å and condensed the beam to \textasciitilde{} 900 nm
diameter, resulting in the beam being completely visible on the camera. To
establish fringe-free conditions, the ``Fine eucentric'' procedure of SerialEM {[}\protect\hyperlink{ref-19ZFerhph}{13}{]} was
used to move a square of the cross-grating grid to the eucentric position of the
microscope. The effective defocus was then set to 2 µm, using the ``autofocus''
routine of SerialEM. The objective focus of the microscope was changed until no
fringes were visible. The stage was then moved in Z until images had an apparent
defocus of 2 µm. The difference in stage Z-position between the eucentric and
fringe-free conditions was used to move other areas into fringe-free condition.

@@

Motion-correction, dose weighting and other preprocessing as detailed below was performed using \emph{cis}TEM {[}\protect\hyperlink{ref-MUY42yxp}{14}{]}. To avoid influence of the beam-edge on motion-correction, only a quarter of the
movie in the center of the camera was considered for calculation of the
estimated motion. After movie frames were aligned and summed, a mask for the
illuminated area was calculated by lowpass filtering the image with a 100 Å
resolution cutoff, thresholding the image at 10\% of the maximal value and then lowpass
filtering the mask again with a 100 Å resolution cutoff to smooth the mask edges. This mask was
then used to select dark areas in the image and fill the pixels with
Gaussian noise, with the same mean and standard deviation as the illuminated
area. A custom version of the unblur program {[}\protect\hyperlink{ref-1G4Y94qCy}{15}{]}
implementing this procedure is available at {[}link to decolace branch{]}. During motion correction images were resampled to a pixel size of 1.5 Å. The
contrast-transfer function (CTF) was estimated using ctffind
{[}\protect\hyperlink{ref-n43f3Sqs}{16}{]}, searching between 0.2 and 2 μm defocus.

@@

The search template was generated from the atomic model of the mouse LSU
(PDB 6SWA, exluding the Epb1 subunit) using the cryo-EM simulator implemented in \emph{cis}TEM
{[}\protect\hyperlink{ref-q2PSamkS}{17}{]}. The
match\_template program {[}\protect\hyperlink{ref-10bXZuF3G}{9}{]} was used to search for this
template in the movie-aligned, exposure-filtered and masked images, using a 1.5°
angular step in out-of-plane angles and a 1.0° angular step in-plane. 11 defocus
planes in 20 nm steps centered around the ctffind-determined defocus were searched. Targets
were defined as detected when their matches with the template produced peaks
with a singal-to-noise ratio (SNR) above a threshold of 7.75, which was chosen
based on the one-false-positive-per-tile criterion {[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]}.

@@

A list of tile pairs \(i,j\) that overlap were assembled by selecting images where
\(|\mathbf{c}_i-\mathbf{c}_j| < D_{Beam}\). In order to calculate the precise offset between tiles \(i\) and \(j\), \(\mathbf{r}_{i,j}\), we calculated the cross-correlation between the two tiles, masked to the overlapping illuminated area using the scikit-image
package {[}\protect\hyperlink{ref-stvWEJeu}{18}{]} was used to calculate refined offsets . The coordinates \(\mathbf{c}_{i}\) were then refined by a least-square minimization against \(\mathbf{r}_{i,j}\):

@@

using the scipy package {[}\protect\hyperlink{ref-8Miti2Gz}{19}{]}. The masked cross-correlation and the least-square minimization was repeated once more to arrive at the final tile alignment.

@@

Images were rendered using UCSF ChimeraX {[}\protect\hyperlink{ref-cYCuuZxG}{20}{]}. The Python scripts used for data processing are available
under {[}repolink{]}.

We used an atomic model of the 60S mouse ribosomal subunit (6SWA) for 2DTM {[}\protect\hyperlink{ref-sbkam02a}{21}{]}. In a
subset of images, the distribution of cross-correlation scores significantly
exceeded the distribution expected from images devoid of detectable targets. In the
resulting scaled maximum-intensity projections (MIPs), clear peaks with SNR
values up to 10 were apparent (Fig. {[}\ref{fig:initmatching2}{]}A). Using a
threshold criterion to select significant targets (see Methods), we found that in
images of cytosolic compartments there were 10-500 ribosomes within one micrograph
(Fig. {[}\ref{fig:initmatching}{]}B-E). Notably, we found no targets in areas
corresponding to the nucleus (Fig. {[}\ref{fig:initmatching}{]}B) or mitochondria
(Fig. \protect\hyperlink{fig:initmatching}{1}D). In the cytoplasm, we found a highly variable
number of targets, only \textasciitilde{} 50 in some exposures (Fig. {[}\ref{fig:initmatching}{]}E) and
up to 500 in others (Fig. {[}\ref{fig:initmatching}{]}C). This is a ten-fold
difference in local ribosome concentration, highlighting the importance of
imaging larger areas of a lamella to gain a complete picture of target
distributions. We therefore set out to collect cryo-EM data for 2DTM from
mammalian cell lamellae in a high-throughput unbiased fashion.

We used this strategy to collect data on eight lamellae, four using the
eucentric focus condition, hereafter referred to as Lamella\textsubscript{EUC}, and four using
the fringe-free condition, hereafter referred to as Lamella\textsubscript{FFF}(Fig.
{[}\ref{fig:assembly}{]} A+D, Fig.
{[}\ref{fig:lamella_images}{]}A). We were able to collect data with a highly
consistent defocus of 800 nm (Fig. {[}\ref{fig:approach}{]}F), both in the eucentric
focus and fringe-free focus condition. To ensure that data were collected
consistently, we mapped defocus values as a function of the applied image shift
(Fig. {[}\ref{fig:lamella_spatial_info}{]}A). This demonstrated that the defocus was
consistent across a lamella, except for rare outliers and in images containing
contamination. We also plotted the measured objective astigmatism of each
lamella and found that it varies with the applied image shift, becoming more
astigmatic mostly due to image shift in the x direction (Fig.
{[}\ref{fig:lamella_spatial_info}{]}B). While approaches exist to correct for this
during the data collection {[}\protect\hyperlink{ref-APSL9LmU}{22}{]}, we opted to not use these approaches in our
initial experiments. We reasoned that because 2DTM depends on high-resolution
information, this would be an excellent test of how much these aberration affect
imaging.

The montages show membrane vesicles and granules with highly variable sizes and
density. We found that a substantial number of granules, which are characterized by higher
density inside the the surrounding cytosol {[}\protect\hyperlink{ref-xE2EhAcH}{23}{]}, seemed to contain a membrane-enclosed inclusion with density similar to
the surrounding cytosol (Fig. {[}\ref{fig:lamella_images}{]}C) and could therefore be formed
by an autophagy-like pathway. These granules were
150-300 nm in diameter and the inclusions were 100-200 nm in diameter. Based on
these dimensions the granules are either azurophil or specific granules {[}\protect\hyperlink{ref-xE2EhAcH}{23}{]}. To our
knowledge, these inclusion have not been described in granulocytes and are
further described and discussed below.

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
%
\documentclass[
]{article}
\usepackage{amsmath,amssymb}
\usepackage{lmodern}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\IfFileExists{bookmark.sty}{\usepackage{bookmark}}{\usepackage{hyperref}}
\hypersetup{
  pdftitle={Defocus Corrected Large Area Cryo-EM (DeCo-LACE) for Label-Free Detection of Molecules across Entire Cell Sections},
  pdfauthor={Johannes ElferichGiulia SchiroliDavid ScaddenNikolaus Grigorieff},
  pdflang={en-US},
  pdfkeywords={cryo-EM, visual protemics, ribosome},
  hidelinks,
  pdfcreator={LaTeX via pandoc}}
\urlstyle{same} % disable monospaced font for URLs
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}
\ifLuaTeX
\usepackage[bidi=basic]{babel}
\else
\usepackage[bidi=default]{babel}
\fi
\babelprovide[main,import]{american}
% get rid of language-specific shorthands (see #6817):
\let\LanguageShortHands\languageshorthands
\def\languageshorthands#1{}
\usepackage[margin=2cm]{geometry}
\usepackage{setspace}
\doublespacing
\usepackage{lineno}
\linenumbers
\usepackage{cleveref}

@@

This manuscript
(\href{https://jojoelfe.github.io/deco_lace_template_matching_manuscript/v/26bf17f811e80363e8f38656b89bbae2c14147f8/}{permalink})
was automatically generated
from \href{https://github.com/jojoelfe/deco_lace_template_matching_manuscript/tree/26bf17f811e80363e8f38656b89bbae2c14147f8}{jojoelfe/deco\_lace\_template\_matching\_manuscript@26bf17f}
on May 27, 2022.

@@

\begin{itemize}
\item
  \textbf{Johannes Elferich}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-9911-706X}{0000-0002-9911-706X}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.pdf}
  \href{https://github.com/jojoelfe}{jojoelfe}
  RNA Therapeutic Institute, UMass Chan Medical School; HHMI
\item
  \textbf{Giulia Schiroli}
  None
\item
  \textbf{David Scadden}
  None
\item
  \textbf{Nikolaus Grigorieff}
  \includegraphics[width=0.16667in,height=0.16667in]{images/orcid.pdf}
  \href{https://orcid.org/0000-0002-1506-909X}{0000-0002-1506-909X}
  · \includegraphics[width=0.16667in,height=0.16667in]{images/github.pdf}
  \href{https://github.com/nikogrigorieff}{nikogrigorieff}
  RNA Therapeutic Institute, UMass Chan Medical School; HHMI
\end{itemize}

@@

We used an atomic model of the 60S mouse ribosomal subunit (6SWA) for 2DTM {[}\protect\hyperlink{ref-sbkam02a}{13}{]}. In a
subset of images, the distribution of cross-correlation scores significantly
exceeded the distribution expected from images devoid of detectable targets. In the
resulting scaled maximum-intensity projections (MIPs), clear peaks with SNR
values up to 10 were apparent (Fig. {[}\ref{fig:initmatching2}{]}A). Using a
threshold criterion to select significant targets (see Methods), we found that in
images of cytosolic compartments there were 10-500 ribosomes within one micrograph
(Fig. {[}\ref{fig:initmatching}{]}B-E). Notably, we found no targets in areas
corresponding to the nucleus (Fig. {[}\ref{fig:initmatching}{]}B) or mitochondria
(Fig. \protect\hyperlink{fig:initmatching}{1}D). In the cytoplasm, we found a highly variable
number of targets, only \textasciitilde{} 50 in some exposures (Fig. {[}\ref{fig:initmatching}{]}E) and
up to 500 in others (Fig. {[}\ref{fig:initmatching}{]}C). This is a ten-fold
difference in local ribosome concentration, highlighting the importance of
imaging larger areas of a lamella to gain a complete picture of target
distributions. We therefore set out to collect cryo-EM data for 2DTM from
mammalian cell lamellae in a high-throughput unbiased fashion.

We used this strategy to collect data on eight lamellae, four using the
eucentric focus condition, hereafter referred to as Lamella\textsubscript{EUC}, and four using
the fringe-free condition, hereafter referred to as Lamella\textsubscript{FFF}(Fig.
{[}\ref{fig:assembly}{]} A+D, Fig.
{[}\ref{fig:lamella_images}{]}A). We were able to collect data with a highly
consistent defocus of 800 nm (Fig. {[}\ref{fig:approach}{]}F), both in the eucentric
focus and fringe-free focus condition. To ensure that data were collected
consistently, we mapped defocus values as a function of the applied image shift
(Fig. {[}\ref{fig:lamella_spatial_info}{]}A). This demonstrated that the defocus was
consistent across a lamella, except for rare outliers and in images containing
contamination. We also plotted the measured objective astigmatism of each
lamella and found that it varies with the applied image shift, becoming more
astigmatic mostly due to image shift in the x direction (Fig.
{[}\ref{fig:lamella_spatial_info}{]}B). While approaches exist to correct for this
during the data collection {[}\protect\hyperlink{ref-APSL9LmU}{14}{]}, we opted to not use these approaches in our
initial experiments. We reasoned that because 2DTM depends on high-resolution
information, this would be an excellent test of how much these aberration affect
imaging.

The montages show membrane vesicles and granules with highly variable sizes and
density. We found that a substantial number of granules, which are characterized by higher
density inside the the surrounding cytosol {[}\protect\hyperlink{ref-xE2EhAcH}{15}{]}, seemed to contain a membrane-enclosed inclusion with density similar to
the surrounding cytosol (Fig. {[}\ref{fig:lamella_images}{]}C) and could therefore be formed
by an autophagy-like pathway. These granules were
150-300 nm in diameter and the inclusions were 100-200 nm in diameter. Based on
these dimensions the granules are either azurophil or specific granules {[}\protect\hyperlink{ref-xE2EhAcH}{15}{]}. To our
knowledge, these inclusion have not been described in granulocytes and are
further described and discussed below.

We used the refined tile positions to calculate the positions of the detected
LSUs in the lamellae (Fig. {[}\ref{fig:matching_euc}{]}A, Fig. {[}\ref{fig:matching_fff}{]}A). Overlaying these positions of the lamellae montages reveal ribosome
distribution throughout the FIB-milled slices of individual cells. Organelles
like the nucleus and mitochondria only showed sporadic targets detected with low
SNRs, consistent with the estimated false-positive rate of one per tile. For
each detected target we also calculated the Z positions from the individual
estimated defocus and defocus offset for each tile. When viewed from the side,
the ribosome positions therefore show the slight tilts of the lamellae relative
to the microscope frame of reference (Fig. {[}\ref{fig:matching_euc}{]}B, Fig.
{[}\ref{fig:matching_fff}{]}B). Furthermore, the side views indicated that lamellae
were thinner at the leading edge. Indeed, when plotting the transmitted beam
intensities in individual tiles as a function of beam image-shift, we observed
substantially higher intensities at the leading edge (Fig. {[}\ref{fig:lamella_motion_thickness}{]}B), which in energy-filtered
TEM indicates a thinner sample {[}\protect\hyperlink{ref-17IP4Lhz1}{16}{]}. Even though we prepared the lamellae
with the ``overtilt'' approach {[}\protect\hyperlink{ref-hr3dyUeI}{17}{]}, this means that ribosome densities across
the lamellae can be skewed by a change in thickness, and better sample
preparation methods are needed to generate more even samples.

Close inspection of the ribosome positions in the lamellae revealed several
interesting features. Ribosomes could be seen associating with membranes, in
patterns reminiscent of the rough endoplasmic reticulum (Fig.
{[}\ref{fig:matching_euc}{]}C, Fig. {[}\ref{fig:matching_fff}{]}C) or the outer nuclear
membrane (Fig. {[}\ref{fig:matching_euc}{]}D). We also observed ribosomes forming
ring-like structures (Fig. {[}\ref{fig:matching_euc}{]}E), potentially indicating
circularized mRNAs {[}\protect\hyperlink{ref-cie13Q8F}{18}{]}. While ribosomes were for the most part excluded
from the numerous granules observed in the cytoplasm, in some cases we observed
clusters of ribosomes in the inclusions of double-membraned granules
described earlier (Fig. {[}\ref{fig:matching_euc}{]}F, Fig. {[}\ref{fig:matching_fff}{]}D,E). It is, in principle, possible that these targets are situated above or
below the imaged granules, since the granule positions in z cannot be determined
using 2D projections. However, in the case of Fig. {[}\ref{fig:matching_fff}{]}E, the
detected ribosomes span the whole lamella in the z direction (Fig. {[}\ref{fig:matching_fff}{]}F), while positions
above or below a granule would result in ribosomes situated exlusively at the
top or bottom of the lamella. This conclusive evidence of ribosomes in
inclusions is consistent with the earlier hypothesis that the inclusions are of
cytoplasmic origin.

We also wondered whether large image shifts would lead aberration due to
astigmatism or beam tilt {[}\protect\hyperlink{ref-APSL9LmU}{14}{]}. We reasoned that if
that was the case the number of detected targets should be highest in the center of the
lamella where the applied beam image-shift is 0. Instead we observed that in
both eucentric and fring-free focus conditions more target where detected at the
``back'' edge of the lamella (Fig. {[}\ref{fig:matching_stat}{]}E{]}). This may be due to the center of the cell being predominantly
occupied by the nucleus, despite its segmentation in neutrophil-like cells. The
increase in matches at the ``back'' of the lamellae compared to the ``front'' can also
be explained by the thickness gradient of the lamellae (Fig. {[}\ref{fig:lamella_motion_thickness}{]}B,
Fig. {[}\ref{fig:matching_euc}{]}B, Fig. {[}\ref{fig:matching_fff}{]}B). In addition, aberrations would be expected to cause average 2DTM SNRs to be higher when beam-image shift
values are small. Instead, we found that SNRs where on average the highes at the
``front'' edge of the lamellae, presumably due to the thinner sample. We therefore
conclude that factors other that beam image-shift or beam
condensation aberrations are limiting 2DTM SNRS, predominantly the thickness of
the lamellae.

In this study we developed an approach to image entire cellular cross-section using cryo-EM at high enough resolution to allow for 2DTM detection of the LSU. The two main advantages compared to previous approaches are a high throughput of imaging and the biological context for detected molecules. The requirement to increase throughput in cryo-EM data collection of cellular samples has been recognized in the recent literature. Most approaches describeb so far are tailored towards tomography. Peck et al. {[}\protect\hyperlink{ref-1GPPSX111}{19}{]} and Yang et al. {[}\protect\hyperlink{ref-ujizcpVn}{20}{]} developed approaches to increase the FOV of tomogram data-collection by using a montaging technique. Peck et al.~used a similar ``condensed-beam'' approach as described here. However, the montages are substantially smaller in scope, covering carbon film holes of 2 um diameter. Bouvette et al. {[}\protect\hyperlink{ref-1Gwx5ScUa}{21}{]} and Eisenstein et al. {[}\protect\hyperlink{ref-xvouNMBy}{22}{]} are using beam image-shift to collect tilt-series in multiple locations in parallel to increase throughput. However, none of these approaches provide the full coverage of a cellular cross-section that can be achieved using DeCo-Lace.

Since we observed substantial variation in ribosome density within and between lamellae, visual proteomics studies that use cryo-EM to establish changes in molecular organization within cells will require orders of magnitude more data than used in this study. One milestone would be to image enough data to represent one cellular volume, which for a small eukaryotic cells requires imaging approximately 100 lamella. While data collection throughput on the TEM is fundamentally limited by the exposure time, this amount of data could be collected within 12 hour by improving the data acquisition scheme to perform all necessary calculation in parallel with actual exposure of the camera. Sample preparation using a FIB/SEM is also currently a bottleneck, but preparation of large lamellae with multiple cellular cross-sections using methods like WAFFLE {[}\protect\hyperlink{ref-17H2LWiIu}{23}{]} might allow sufficient throughput. As stated in the results, at least for 2DTM computation will remain challenging and approximately 17,000 GPU hours would be required for a 100 lamellae dataset.

As described in {[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]} the 2DTM SNR threshold for detecting a target is chosen to result in one false positive detection per image searched. We would therefore expect to find one false positive detection per tile. We reasoned that the large nuclear area imaged by DeCo-Lace could be used to test whether this assumption is true. In the 670 tiles containing exclusively nucleus (as manually annotated from the overview image) we detected 247 targets, making the false-positive rate more than twofold lower than expected. Since earlier work shows that 2DTM with the LSU can produce matches to nuclear ribosome biogenesis intermediates {[}\protect\hyperlink{ref-T1Qr0xyP}{24}{]}, this could even be an overestimate of the false-positive rate. This suggests that the detection threshold could be even lower, which is an area of ongoing research.

We found that even though we used beam image-shift extensively (up to 7 um), we did not see substantially reduced 2DTM SNR values in tiles acquired at high beam image-shift compared to tiles acquired with low or no beam image-shift. This is in contrast to reports in single-particle analysis (SPA) {[}\protect\hyperlink{ref-44fG6kcy}{25}{]} where the induced beam tilt substantially reduced the resolution if it was not corrected during processing. It is possible that 2DTM is less sensitive to beam-tilt aberrations, since the template is free of any aberration and only the image is distorted, while in SPA the beam tilt will affect both the images and the reconstructed template.

Unexpectedly, we observed granules containing a vesicle of putative cytosolic origin. We speculate that upon degranulation, the process in which granules fuse with the plasma membrane, these vesicles would be released into the extracelullar space. The main types of extracellular vesicles of this size are exosomes, up to 100 nm large vesicles derived from fusion of multivesicular bodies with the plasma membrane, and microvesicles, which are derived from direct budding of the plasma membrane {[}\protect\hyperlink{ref-kncnYGLD}{26}{]}. We suggest that granulocytes could release a third type of extracellular vesicle, granule-derived vesicles (GDV), into the extracellular space. 2DTM showed that a subset of GDVs can contain ribosomes (Fig. {[}\ref{fig:matching_euc}{]}F, Fig. {[}\ref{fig:matching_fff}{]}D,E). This could indicate that these vesicles are transporting translation-capable mRNAs, as has been described for exosomes {[}\protect\hyperlink{ref-agVdpX76}{27}{]}. Further studies will be necessary to confirm the existence of GDVs in granulocytes isolated from mammals and to understand their functional significance.

@@

Grids were loaded into a Titan Krios TEM (Thermo Fisher) operated at 300 keV and
equipped with a BioQuantum energy filter (Gatan) and K3 camera (Gatan). The
microscope was aligned using a cross-grating grid on the stage. Prior to each
session, we carefully performed the ``Image/Beam'' calibration in nanoprobe. We
set the magnification to a pixel size of 1.76 Å and condensed the beam to \textasciitilde{} 900 nm
diameter, resulting in the beam being completely visible on the camera. To
establish fringe-free conditions, the ``Fine eucentric'' procedure of SerialEM {[}\protect\hyperlink{ref-19ZFerhph}{28}{]} was
used to move a square of the cross-grating grid to the eucentric position of the
microscope. The effective defocus was then set to 2 µm, using the ``autofocus''
routine of SerialEM. The objective focus of the microscope was changed until no
fringes were visible. The stage was then moved in Z until images had an apparent
defocus of 2 µm. The difference in stage Z-position between the eucentric and
fringe-free conditions was used to move other areas into fringe-free condition.

@@

Motion-correction, dose weighting and other preprocessing as detailed below was performed using \emph{cis}TEM {[}\protect\hyperlink{ref-MUY42yxp}{29}{]}. To avoid influence of the beam-edge on motion-correction, only a quarter of the
movie in the center of the camera was considered for calculation of the
estimated motion. After movie frames were aligned and summed, a mask for the
illuminated area was calculated by lowpass filtering the image with a 100 Å
resolution cutoff, thresholding the image at 10\% of the maximal value and then lowpass
filtering the mask again with a 100 Å resolution cutoff to smooth the mask edges. This mask was
then used to select dark areas in the image and fill the pixels with
Gaussian noise, with the same mean and standard deviation as the illuminated
area. A custom version of the unblur program {[}\protect\hyperlink{ref-1G4Y94qCy}{30}{]}
implementing this procedure is available at {[}link to decolace branch{]}. During motion correction images were resampled to a pixel size of 1.5 Å. The
contrast-transfer function (CTF) was estimated using ctffind
{[}\protect\hyperlink{ref-n43f3Sqs}{31}{]}, searching between 0.2 and 2 μm defocus.

@@

The search template was generated from the atomic model of the mouse LSU
(PDB 6SWA, exluding the Epb1 subunit) using the cryo-EM simulator implemented in \emph{cis}TEM
{[}\protect\hyperlink{ref-q2PSamkS}{32}{]}. The
match\_template program {[}\protect\hyperlink{ref-10bXZuF3G}{9}{]} was used to search for this
template in the movie-aligned, exposure-filtered and masked images, using a 1.5°
angular step in out-of-plane angles and a 1.0° angular step in-plane. 11 defocus
planes in 20 nm steps centered around the ctffind-determined defocus were searched. Targets
were defined as detected when their matches with the template produced peaks
with a singal-to-noise ratio (SNR) above a threshold of 7.75, which was chosen
based on the one-false-positive-per-tile criterion {[}\protect\hyperlink{ref-Ynb3IP6I}{7}{]}.

@@

A list of tile pairs \(i,j\) that overlap were assembled by selecting images where
\(|\mathbf{c}_i-\mathbf{c}_j| < D_{Beam}\). In order to calculate the precise offset between tiles \(i\) and \(j\), \(\mathbf{r}_{i,j}\), we calculated the cross-correlation between the two tiles, masked to the overlapping illuminated area using the scikit-image
package {[}\protect\hyperlink{ref-stvWEJeu}{33}{]} was used to calculate refined offsets . The coordinates \(\mathbf{c}_{i}\) were then refined by a least-square minimization against \(\mathbf{r}_{i,j}\):

@@

using the scipy package {[}\protect\hyperlink{ref-8Miti2Gz}{34}{]}. The masked cross-correlation and the least-square minimization was repeated once more to arrive at the final tile alignment.

@@

Images were rendered using UCSF ChimeraX {[}\protect\hyperlink{ref-cYCuuZxG}{35}{]} using a custom extension to render 2DTM results available at {[}https://github.com/jojoelfe/tempest{]}. The Python scripts used for data processing are available
under {[}https://github.com/jojoelfe/deco\_lace\_template\_matching\_manuscript{]}.

\hypertarget{acknowledgments}{%
\subsection{Acknowledgments}\label{acknowledgments}}

The authors would like to thank Bronwyn Lucas, Carsten Sachse, and Chen Xue for helpful suggestions and careful reading of the manuscript. Data was collected at the UMass Chan medical School CryoEM core with help by Kankang song, Christna, and Chen Xue.

\hypertarget{data-availability}{%
\subsection{Data availability}\label{data-availability}}

Cryo-EM movies, motion-corrected images and template matching results have been deposited in EMPIAR under accession code {[}CODE{]}. The custom cisTEM version is available here {[}{]} until features have been integrated into a cisTEM release. The ChimeraX extension for rendering is avaibale here {[}{]}. This manuscript was prepared using the manubot package {[}@ cite{]} . Custom python scripts and all files nescessary for manuscript generation are available here {[}{]}.
